Object-based Auditory Scenes Similarity Retrieval and Classification With Wavelet Scattering

### 2016-07-11 paper/structureScene.tex
update EMD description and object based approach description

The proposed approach is motivated in Section \ref{sec:motivations} and a brief review the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}. Then, the contributions are extensively described in Section \ref{sec:scattering} and \ref{sec:object}, respectively. Experiments carried out in an unsupervised setting, \ie acoustic scene similarity (ASS) and in a supervised setting, \ie acoustic scene classification (ASC) are subsequently reported in Section \ref{sec:experiments}.

@@

Although the precise set of events occuring within a given auditory stream may not be discernable, even for trained humans, research shows that the focus on a few events, so called markers,  sufficed to reliably predict many high-level attributes \cite{} \vl{ref}.
For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas \gl{citations ?}.
\vl{Missing transition}

@@

\subsection{Logarithmic compression}
Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances.
Yet, because of the complex modulus nonlinearity, scattering coefficients are nonnegative by design.
It appears empirically that their distribution is skewed towards the right, which means that the tail towards greater values is longer than the tail towards lower values.
However, skewness can be reduced by applying to all coefficients a pointwise concave transformation, \eg logarithmic.
Figure \ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE 2013 dataset, before and after logarithmic compression.

@@

\section{Object-based modeling of the scene}

It has been shown the perception of complex acoustic scene rely on the identification of putative sound sources.
\ml{GL 2 colonnes}

describe in terms of high level attributes

be they cognitive ones such as pleasantness, or descriptive ones such as the type of environment where the

@@

Whilst the range of applications are very large, so is the need

holistic / event based

skeleton of events on a bed of texture.

debate about the kind of process underlying perception

strong indication that the detection of specific events, named markers suffice to trigger class prediction

From a numerical data processing overview, the holistic scheme as a simplicity, but clearly face poor performance on realistic conditions \cite{lagrange:hal-01082501}.

If available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas.

To take into account this object-based representation, we propose the following 4-steps algorithm:

\begin{itemize}
\item step 1: features extraction.
\item step 2: scene-wise clustering step. A clustering is performed on each scene using an enhanced version of the popular K-means algorithm known as K-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points. Each scene is then describe with a set of centroids.
\item step 3: Centroids similarity. The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}: \\
\begin{equation}
\label{eq:kc}
\begin{split}
K^c_{ij} & =exp\left( - \dfrac{\parallel c_i - c_j \parallel^2}{\sigma_{i}\sigma_{j}} \right) \\
\sigma_{i} & =\parallel c_i - c_N \parallel
\end{split}
\end{equation}

@@

\item step 4: scene similarity \\
\end{itemize}

The scene similarity is computed based on the extracted centroids. We propose the three following strategies:

\begin{itemize}
\item \emph{closest}: the similarity between two scenes is equal to the largest similarity between their centroids.
\item \emph{averaged}: the similarity between two scenes is equal to the average of their centroids similarities.
\item \emph{weighted}: for each scene, each centroid is weighted according to the number of frames belonging to its cluster.\\
\end{itemize}

Compared to the closest or averaged approach, the  weighted similarity between two scenes takes into account both the clusters weights and the centroids similarities. To do so, each scene $s_u$ is viewed as a signature $S_u$ of $m$ clusters ($S_u=\lbrace(c_1^u,w_1^u),\ldots,(c_m^u,w_m^u)\rbrace$), with $c_i^u$ and $w_i^u$ being the centroid and the weight of the $i$th cluster of the scene $s_u$ respectively.

To compute the distance between two signatures $S_u$ and $S_v$, we use a variant version of the Earth Moving Distance ($EMD$), known as $\widehat{EMD}$ introduced in \cite{pele2008linear}. $EMD$ is a cross-bin histogram distance, widely used in computer vision (\cite{zhang2007local}), which compute the minimal cost to be paid to transform one histogram into the other. The advantage of using the $EMD$ is that it does not require bins to be aligned. The alignment problem is solved by using a ``\,ground distance\,'', being the distance between the bins representatives of the two histograms. In our case, the histograms are the weights $w$, and the bins representative the centroids $c$.\\

@@

where $f_{ij}$ is the optimal flow between the clusters weights $w_i^u$ and $w_j^v$, that is, the amount transported from the $i$th bin to the $j$th bin. $D_{ij}$ is the ground distance between the centroids $c_i^u$ and $c_j^v$, which is equal to $D_{ij}=1-K^c_{ij}$.\\

@@

\begin{equation*}
K_{uv}^s(s_u,s_v) = exp\left( - \dfrac{\widehat{EMD}(S_u,S_v)}{A} \right) \\
\end{equation*}

@@

@@

The influence of the temporal-scattering transform and the object-based approach (\emph{ob}) are assessed in an unsupervised setting (\emph{i.e.} ASS) and in a supervised setting (\emph{i.e.} ASC).

For the ASS setting, evaluation is performed on the private dataset of the DCASE 2013 challenge. The metric used is the precision at rank $k$ ($p@k$), which is the  average  number  of  items  of  the  same  class among the $k$ closest items averaged for all possible seed items. The $p@k$ is computed for $k=[1,\ldots,9]$. Note that a $p@1$ is equivalent to an accuracy computed based on the labels of the 1st nearest neighbors of the items to classify.

For the ASC setting, the systems are evaluated  with five-fold  stratified cross validation on the private dataset of the DCASE 2013 challenge. The folds are identical to those used in the original challenge. Optimal parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approach and the scaling parameters $N$ of the RBF kernels (see Eq. \ref{eq:kc}), are learned on the public dataset of the DCASE 2013 challenge, using five-fold  stratified cross validation as well.

\subsection{Comparison with early and late integration}

@@

Experiments are carried out using temporal-scattering coefficients as well as Mel Frequency Cesptral Coefficients (MFCCs) as features. For the temporal-scattering, each scene is described by 128 vectors of scattering coefficients \gl{vincent: windows and hops ?}. Different sets of parameters were tested to compute the MFFCs. We here report only the results for the best setting, which is similar to the one used by the winners of the previous DCASE 2013 challenge \cite{roma2013}. 40 MFCCs (including the first energy coefficient) are computed for windows of 50ms and hops of 25ms, with a limited frequency range going from $27.5$Hz to $1000$Hz. This setting gives us 600 feature vectors per scenes. It has been reported that classifying the full set of frames directly may not be the best approach. In order to reduce the data dimensionality, a pooling step is performed on the MFCCs vectors: each set of MFCCs vectors is dividing into $n$ non-overlapping $t$-seconds long slices. Each $n$ MFCC slices is then averaged over time. As the \emph{ob} approach involves a clustering step, it is desirable to ensure that each scene is described by circa the same number of vectors,for both the temporal-scattering and the MFCCs, prior to running the clustering. Thus, $t$ is set to 250ms, which gives us 120 MFCCs vectors per scenes.

@@

As most of the 2013 DCASE challenge participant did, a Support Vector Machine (SVM), which is a large-margin binary discriminators with tunable regularization of misclassified examples, is used as classifier for the ASC setting. For the \emph{ob} approach, we use XX. For both the \emph{ei} and \emph{li} approaches, a linear kernel as well as a RBF kernel are used. As for the \emph{ob} approach, the RBF kernel is scaled locally using the method proposed in \cite{selfTuneManor2004}. All features, that is the frames, the pooled frames or the centroids depending on the approach considered, are standardized prior to be filled into the classifier. The standardization is computed with respect to the train/test set split of the cross validation scheme (see \ref{sec:stand}).

>>>>

The proposed approach is motivated in Section \ref{sec:motivations} and a brief review the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}. Then, the contributions are extensively described in Section \ref{sec:scattering} and \ref{sec:object}, respectively. Experiments carried out in an unsupervised setting, \ie acoustic scene similarity (ASS) and in a supervised setting, \ie acoustic scene classification (ASC) are subsequently reported in Section \ref{sec:ExpEval}.

@@

Although the precise set of events occuring within a given auditory stream may not be discernable, even for trained humans, research shows that the focus on a few events, so called markers,  sufficed to reliably predict many high-level attributes \cite{lafayPartI}.
For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas \gl{citations ?}.
\vl{Missing transition}

@@

\subsection{Logarithmic compression}
\label{sec:logcomp}

Many algorithms in pattern recognition, including nearest neighbor classifiers and support vector machines, tend to work best when all features follow a standard normal distribution across all training instances.
Yet, because of the complex modulus nonlinearity, scattering coefficients are nonnegative by design.
It appears empirically that their distribution is skewed towards the right, which means that the tail towards greater values is longer than the tail towards lower values.
However, skewness can be reduced by applying to all coefficients a pointwise concave transformation, \eg logarithmic.
Figure \ref{fig:histograms} shows the distribution of an arbitrarily chosen scattering coefficient over the DCASE 2013 dataset, before and after logarithmic compression.

@@

\section{Object-based modeling of the scene}
\label{sec:object}

@@

Studies in cognitive psychology investigating urban sound environment quality pointed out that low level holistic descriptors (perceived or measured) such as global sound level, are not sufficient to fully characterized an acoustic scene \cite{guyot2005urban,kang2006urban}. Instead several papers have reported that some cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness jugdment \cite{kuwano_memory_2003}  relies upon high level attributes such as the identification and assessment of the sound source which composed the scene. It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas \cite{lafayPartI}. Thus recent researches in sound perception now focus on the study of the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}, viewing an acoustic scene as ``\,skeleton of events on a bed of texture\,''  \cite{nelken_ear_2013}. \gl{debate about the kind of process underlying perception: holistic / event based perception ?}. \\

From a numerical data processing point of view, a holistic scheme, such as the bag-of-frame approach \cite{aucouturier2007bag}, has a simplicity, but clearly face poor performance on realistic conditions \cite{lagrange:hal-01082501}. Instead the outcomes in sound perception cited above encourage us to adopt an object-based  representation of the scene. \gl{vincent, un petit mot pour motiver le fait que le scattering est un bon outil pour separer les evenements peuplant une scène sonore ?}. \\

In a first attempt to make use of an object-based (\emph{ob}) representation of the scene, we propose the follow a clustering approach. After a features extraction step,  a clustering is performed on each scene features using an enhanced version of the popular K-means algorithm known as K-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points. \\

Each scene is then described with a set of centroids. To compute the similarities between the scene based on their extracted centroids, we propose three strategies:

\begin{itemize}
\item \emph{ob-closest}: the similarity between two scenes is equal to the largest similarity between their centroids.
\item \emph{ob-averaged}: the similarity between two scenes is equal to the average of their centroids similarities.
\item \emph{ob-weighted}: for each scene, each centroid is weighted according to the number of frames belonging to its cluster.\\
\end{itemize}

The similarity between two centroids $c_i$ and $c_j$ is computed using a radial basis function kernel $K^c$ combined with the local scaling method proposed in \cite{selfTuneManor2004}: \\
\begin{equation}
\label{eq:kc}
\begin{split}
K^c_{ij} & =exp\left( - \dfrac{\parallel c_i - c_j \parallel^2}{\sigma_{i}\sigma_{j}} \right) \\
\sigma_{i} & =\parallel c_i - c_N \parallel
\end{split}
\end{equation}

@@

Compared to the closest or averaged approach, the  weighted similarity between two scenes takes into account both the centroids similarities and the clusters weights. Each scene $s_u$ is described by a signature $S_u$ of $m$ clusters ($S_u=\lbrace(c_1^u,w_1^u),\ldots,(c_m^u,w_m^u)\rbrace$), with $c_i^u$ and $w_i^u$ being the centroid and the weight of the $i$th cluster of the scene $s_u$ respectively.

To compute the distance between two signatures $S_u$ and $S_v$, we use a variant version of the Earth Moving Distance ($EMD$), known as $\widehat{EMD}$ and introduced in \cite{pele2008linear}. $EMD$ is a cross-bin histogram distance, widely used in computer vision \cite{zhang2007local}, which compute the minimal cost to be paid to transform one histogram into the other. The advantage of using the $EMD$ is that it does not require bins to be aligned. The alignment problem is solved by using a ``\,ground distance\,'', which is the distance between the bins representatives of the two histograms. In our case, the histograms are the cluster weights $w$, and the bins representative the cluster centroids $c$.\\

@@

where $\lbrace f_{ij} \rbrace$ is the optimal flow between the cluster weights $w_i^u$ and $w_j^v$, that is, the amount transported from the $i$th bin to ``\,supply to the demand\,'' of the $j$th bin. $D_{ij}$ is the ground distance between the centroids $c_i^u$ and $c_j^v$, and is computed from the kernel $K^c$:

\begin{equation*}
D_{ij}=1-K^{c(uv)}_{ij}
\end{equation*}

with $K^{c(uv)}$ being the slice of $K^{c}$ containing the pairwise similarities between the centroids sets $c^u$ and $c^v$ belonging to the scenes $s_u$ and $s_v$ respectively. As suggested in \cite{pele2009fast}, we set the $\alpha$ parameter to one.\\

@@

\begin{equation}
\label{eq:ks}
K_{uv}^s = exp\left( - \dfrac{\widehat{EMD}(S_u,S_v)}{A} \right) \\
\end{equation}

@@

Experiments are carried out using temporal-scattering coefficients as well as Mel Frequency Cesptral Coefficients (MFCCs) as features. For the temporal-scattering, each scene is described by 128 vectors of scattering coefficients \gl{vincent: windows and hops ?}. Experiments are conducted with and without the log compression (see Section \ref{sec:logcomp}).

Different sets of parameters were tested to compute the MFFCs. We here report only the results for the best setting, which is similar to the one used by the winners of the previous DCASE 2013 challenge \cite{roma2013}. 40 MFCCs (including the first energy coefficient) are computed for windows of 50ms and hops of 25ms, with a limited frequency range going from $27.5$Hz to $1000$Hz. This setting gives us 600 feature vectors per scenes. It has been reported that classifying the full set of frames directly may not be the best approach \cite{rakotomamonjy2015histogram}. In order to reduce the data dimensionality, a pooling step is performed on the MFCCs vectors: each set of MFCCs vectors is dividing into $n$ non-overlapping $t$-seconds long slices. Each $n$ MFCC slices is then averaged over time. As the \emph{ob} approach involves a clustering step, it is desirable to ensure that each scene is described by circa the same number of vectors,for both the temporal-scattering and the MFCCs, prior to running the clustering. Thus, $t$ is set to 250ms, which gives us 120 MFCCs vectors per scenes.

@@

The influence of the temporal-scattering transform and the \emph{ob} are assessed in an unsupervised setting (\emph{i.e.} ASS) and in a supervised setting (\emph{i.e.} ASC).

For the ASS setting, evaluation is performed on the private dataset of the DCASE 2013 challenge. The metric used is the precision at rank $k$ ($p@k$), which is the  average  number  of  items  of  the  same  class among the $k$ closest items averaged for all possible seed items. The $p@k$ is computed for $k=[1,\ldots,9]$. Note that a $p@1$ is equivalent to an accuracy computed based on the labels of the 1st nearest neighbors of the items to classify. For each experimental setting, we report the results for the set of parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approach and the scaling parameters $N$ of the RBF kernels (see Eq. \ref{eq:kc}) giving the best $p@9$.

For the ASC setting, the systems are evaluated  with five-fold  stratified cross validation on the private dataset of the DCASE 2013 challenge. The folds are identical to those used in the original challenge. Optimal parameters (see above) are this time learned on the public dataset of the DCASE 2013 challenge, using five-fold  stratified cross validation as well.

@@

@@

As most of the 2013 DCASE challenge participants did, a Support Vector Machine (SVM), which is a large-margin binary discriminators with tunable regularization of misclassified examples, is used as classifier for the ASC setting. The parameter $C$ is here set to one.

For the \emph{ob} approaches (\emph{weighted}, \emph{closest} and \emph{average}), we use the precomputed kernels described in Section \ref{sec:object}. For both the \emph{ei} and \emph{li} approaches, a linear kernel as well as a RBF kernel are used. As for the \emph{ob} approaches, the RBF kernel is scaled locally using the method proposed in \cite{selfTuneManor2004}. All feature vectors, being frames, pooled frames or centroids, depending on the approach considered, are standardized prior to be filled into the classifier. The standardization is computed with respect to the train/test set splits of the cross validation scheme (see Section \ref{sec:stand}).
